---
title: "KiddoPrediction"
author: "Benjamin Harder"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(rstan)
```

## R Markdown

Some good friends of ours are expecting their first child, and like most parents to be, they have started a "guess the characteristics of the baby" betting pool.  

1. Will it be a boy or girl?
2. On what day will the baby be born?
3. What time will the baby be born?
4. How much will the baby weigh? (in pounds/oz)
5. How long will the baby be? (in inches)
6. What will the baby's name be?

This topic is interesting because kids are miracle and a **mystery of the universe.**  From a statistical perspective, I'm going to assume that the attributes surrounding the birth characteristics of a newborn child come from a deeply stochastic process.  Given this assumption of randomness, it begs the question - can data and analysis give me an edge in choosing the right characteristics and winning the bet?  

We do know some demographic characteristics about her that can help guide the analysis:

* Her due date is July 14th, 2020
* She is in her late 30s
* They do not know the gender
* She has had an easy pregrancy with no complications
* She lives in Canada3
* They will be having one child, so no twins
* This is her first child
* She plans on having a natural birth


### Data
```{r cars}
baby_data_all <- read_csv(
    "birthdata.csv",
    col_types = cols(
      motherbirthdate = col_double(),
      duedate = col_double(),
      birthdate = col_double())) %>%
  #select only relevant columns
  select(motherbirthyear, 
         duedate,
         birthdate,
         country, 
         multiples, 
         previousbirths, 
         cesarean,
         sex,
         pounds,
         ounces) %>%
  #coerce unix timestamps into datetimes
  mutate(duedate = as_datetime(duedate),
         birthdate = as_datetime(birthdate)) %>% 
  #filter out irrelevant characteristics
  filter(multiples == 1) %>%  #only one kid
  filter(previousbirths == 0) %>%  #no previous kids
  filter(cesarean == 0) %>%  #natural birth
  filter(sex != "NULL") %>% #filter null 
  mutate(monther_age = 2020 - motherbirthyear, #age of the mother
         miss_date = as.double(difftime(birthdate, #delta from due date
                                        duedate, unit="days")), 
         weight_oz = (pounds * 16) + ounces, #total weight in oz
         sex = as.factor(sex)) %>%  #factorize
  drop_na() %>% #no NAs
  filter(monther_age %in% (30:45)) %>% #age range
  filter(miss_date %in% (-20:20)) %>%  #missing date
  filter(weight_oz > 1) %>%  #bad entries for weight
  select(birthdate, country, sex, miss_date, pounds, ounces, weight_oz)
```
```{r}
#Canadian Filter
#there are a lot of Canadas in this data, need to combine them into one canada lable
baby_data_ca <- baby_data_all %>% 
  filter(country %in% c("Canada", "canada", "CANADA", "CAnada")) %>% 
  select(-country)
```


### Analysis

#### Will it be a boy or girl?
```{r}
table(baby_data_ca$sex)
prop.table(table(baby_data_ca$sex))
```
Well this certainly looks like a crapshoot - maybe a simple binomial test can help determine if there is a difference between the boy/girl split

```{stan output.var="gender_diff", eval=FALSE}
data {
  // Number of data points
  int n1;
  // Number of successes
  int y1[n1];
}

parameters {
  real<lower=0, upper=1> theta;
}

model {  
  theta ~ beta(1, 1);
  y1 ~ bernoulli(theta);
}

generated quantities {}
```


```{r}
#gender distro vector
baby_gender <- c(rep(0, 151), rep(1, 144))
gender_data_list <- list(y1 = baby_gender, n1 = length(baby_gender))
stan_samples <- sampling(gender_diff, data = gender_data_list)
s <- as.data.frame(stan_samples)
```

```{r}
hist(s$theta)
mean(s$theta > 0.5)
```

#### On what day will the baby be born?
```{stan output.var = born_date, eval = False}
data {
  int n1;
  vector[n1] y1;
}
parameters {
  real mu1;
  real<lower=0> sigma1;
}
model {  
  mu1 ~ uniform(-40, 40);
  sigma1 ~ uniform(0, 1000);
  y1 ~ normal(mu1, sigma1);
}
generated quantities {}
```


```{r}
data_list <- list(y1 = baby_data_ca$miss_date,
                  n1 = length(baby_data_ca$miss_date))

stan_samples <- sampling(born_date, data = data_list)
```
```{r}
s <- as.data.frame(stan_samples)
ggplot(s, aes(x = mu1)) +
         geom_histogram(color="black", fill="lightblue") +
  geom_vline(xintercept = mean(s$mu1), color="red", linetype=1, size =2) +
  ggtitle(expression(paste("Posterior estimate for ", mu))) + 
  theme_classic()
```

```{r}

```


#### What time will the baby be born?
```{r}

```

#### How much will the baby weigh? (in pounds/oz)
```{r}

```

#### How long will the baby be? (in inches)
```{r}

```

#### What will the baby's name be?
```{r}

```

### Final Recommendations


### Conclusion/Alternatives



```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
